\section{Mathematical Resolution and Computational Bounds}

Establishing topological invariance is a prerequisite; the subsequent requirement is demonstrating that the generated braid group representations are dense in the special unitary group, thereby achieving universal quantum computation capabilities.

\subsection{Modular Tensor Categories and the Quantum Double}
A Modular Tensor Category (MTC) serves as the complete algebraic ecosystem that governs topological quantum systems. It provides the exact mathematical rules for how anyons combine (fusion) and how they exchange positions (braiding). Formally, an MTC is a braided monoidal category equipped with additional structure: it is ribbon (possessing compatible twists and duals), semi-simple, and possesses a non-degenerate $S$-matrix~\cite{bakalov2001}. 

The UOR Atlas leverages a specific MTC derived from the generalized quantum double $D(G)$, where $G$ is a finite group. Specifically, the framework constructs representations analogous to $D(\mathbb{Z}_O)$~\cite{drinfeld1987}, allowing the 24 classes to algebraically execute the fusion and braiding rules of non-abelian excitations (such as Fibonacci anyons, where fusing two $\tau$ anyons yields either nothing or another $\tau$, expressed as $\tau \otimes \tau = I \oplus \tau$).

\subsection{Density in \texorpdfstring{$SU(2)$}{SU (2)} via Freedman-Larsen-Wang}
For a topological quantum computer to be strictly "universal," it must be able to perform any possible quantum computation. In mathematical terms, the physical braid operations must be capable of smoothly approximating any continuous rotation in the quantum state space (known as the special unitary group, $SU(2^n)$). If the available operations can get arbitrarily close to any target rotation, the set is mathematically \textit{dense}.

To mathematically prove both computational efficiency and universal capability, the framework evaluates the density of the generated subgroups using the Freedman-Larsen-Wang (FLW) theorem~\cite{freedman2002}.

\begin{proof}
Let $\mathcal{G} = \{g_1, g_2, \dots, g_k\}$ represent the fundamental topological generators derived from the invariant symmetry classes. The structural properties of this group dictate its computational power:

\textbf{1. Finite Braid Closure (Modular Data):} The modular subgroup generated by the $S$ and $T$ matrices on the $24$-dimensional space is mathematically finite. This closure was verified computationally via a bounded Breadth-First Search (BFS) exploration over the modular data that successfully terminated. This finite representation mathematically establishes the \textit{cache-collapse advantage}, permitting the system to undergo $O(N^6)$ classical topological verification without intractable state expansion.

\textbf{2.
Density on the 2D Commutant Block: Exactly Decided, Currently Refuted:} The density question for the coupled generators is decided by an exact algebraic certifier over the cyclotomic field $\mathbb{Q}(\zeta_{24})$, replacing the earlier $f64$ procedure that witnessed the non-collapse hypothesis against a numerical threshold ($|\beta|^2 > 10^{-4}$).
The exact computation is unconditional on the algebraic side and uses a single analytic input, the transcendence of $t = e^{i}$ (Lindemann): grading the commutant equations by powers of $t$ reduces the commutant of the coupled generators to an explicit linear system over $\mathbb{Q}(\zeta_{24})$, whose kernel has exact dimension 2, yielding a unique 2-dimensional invariant block with exact projector $P_1$, $\operatorname{tr} P_1 = 2$.
The exact verdict refutes density on this block.
First, $\operatorname{tr}(P_1 G_S) = 0$ identically: every trace coefficient the earlier threshold test reported as non-zero is exactly zero, so the restricted $S$-generator is a projective involution.
Second, the block is supported entirely in the $(-1)$ spectral eigenspace, where the archimedean coupling $E$ restricts to the scalar phase $t^{-1}$; the coupling is therefore projectively trivial on the block.
Third, the projective image of the generators on the block is finite, of exact order 24, the projective single-qubit Clifford group $S_4$, consistent with the abelian character of the modular sector.
The correct density criterion runs in $PU(2) \cong SO(3)$, whose closed subgroups are finite, $SO(2)$, $O(2)$, or $SO(3)$; a certificate requires a projectively non-commuting pair of projectively infinite-order elements, each property decidable as an exact Laurent-polynomial identity over $\mathbb{Q}(\zeta_{24})$.
No such pair exists on this block.
The Solovay-Kitaev density question is accordingly recorded as decided in the V\&V dictionary: the row asserts the exact decision as a theorem, including the refutation, leaving no open question.
The continuous content of the coupled machine is not absent but located: on the 22-dimensional irreducible complement, whose support straddles all four spectral eigenspaces and where the coupling $E$ is non-scalar, the adjoint-trace criterion (if $u$ has finite projective order, $\operatorname{tr}(\operatorname{Ad} u) = \operatorname{tr}(u)\overline{\operatorname{tr}(u)}$ is algebraic; as an exact Laurent polynomial in $t$, a nonzero coefficient at a nonzero grade forces transcendence) certifies infinite projective order for the generator words, and a projectively non-commuting pair is exhibited.
The projective closure on the 22-dimensional block is therefore an infinite non-abelian compact group acting irreducibly: the coupled generators exceed every finite gate set, and the beyond-Clifford content is exactly where the archimedean coupling acts non-trivially.
The identity component of that closure is then determined: by Kronecker-Weyl and the irrationality of $\pi$, the spectral flow $\exp(i\mathbb{R}M)$ lies in the closure of the discrete group, placing $iM$ in its Lie algebra with integer entries; a division-free Lie closure under $\operatorname{Ad}(S)$, $\operatorname{Ad}(T)$ (both honestly in the closure, since $E = \exp(iM)$ lies in the spectral flow), brackets, and the $\operatorname{ad}(iM)$-weight splitting yields a sound mod-$p$ lower bound of 483 on $\dim \operatorname{Lie}(H)$ restricted to the block.
Since $\mathfrak{su}(22)$ is simple with minimal proper-subalgebra codimension $2n-2 = 42$, any subalgebra of $\mathfrak{u}(22)$ of dimension at least 483 contains $\mathfrak{su}(22)$; the projective closure therefore contains $PSU(22)$ and is dense in $PU(22)$.
The coupled Atlas generators natively support universal quantum computation on a 22-dimensional qudit carrier, by Solovay-Kitaev in $PU(d)$.
The density question is thus decided in full, in both directions: refuted on the 2-dimensional block where the coupling is projectively scalar, and established as $PU(22)$-density on the 22-dimensional block where it is not.
The multi-carrier question is likewise decided exactly, in three theorems.
First, a separation theorem: no power of the native monodromy preserves the tensor code formed by two 22-dimensional blocks (exactly, no monodromy power is constant on the component pairs of the block-projector support graph), so the native diagonal sector cannot entangle the continuous carriers and the tensor-code architecture is closed.
Second, irreducibility: the two-handle native group, generated by the per-handle coupled generators and the monodromy, has exact commutant dimension 1 on the 576-dimensional pair space; the monodromy dissolves the block structure, and the correct multi-handle carrier is the irreducible pair block itself.
Third, native continuous entanglement: embedding the per-handle Lie algebras and adjoining the monodromy conjugates yields a sound mod-$p$ lower bound of 986 on $\dim \operatorname{Lie}(H_2)$, exceeding the maximal local subalgebra bound $2 \cdot \dim(\mathfrak{u}(2) \oplus \mathfrak{u}(22)) = 976$: the identity component of the two-handle closure contains non-local directions, so continuous entangling flows exist natively on the irreducible pair carrier.
The pair carrier then saturates.
Since each $\operatorname{End}(\mathbb{C}^{24})$ contains the adjoint of the corner $\mathfrak{su}(22)$ exactly once, $\mathrm{adj}\otimes\mathrm{adj}$ occurs with multiplicity one in $\mathfrak{gl}(576)$ as a module over $\mathfrak{s} = \mathfrak{su}(22)\oplus\mathfrak{su}(22)$; a certified nonzero $\mathrm{adj}\otimes\mathrm{adj}$ component of one element of $\operatorname{Lie}(H_2)$ (nonzero mod $p$, hence nonzero exactly) therefore forces the entire component inside, and bracket closure with corner-supported anticommutators yields the full $\mathfrak{su}(484)$ on $F = W'\otimes W'$.
With $\mathfrak{su}(F)$ inside, the $F$-isotypic part of ${\operatorname{Lie}(H_2)}_{\mathbb{C}}$ in the off-diagonal blocks is $S\otimes F \oplus \bar S\otimes F^*$ for a unique $S\subseteq\mathbb{C}^{92}$, and the certified reachability rank $92$ (the ambient cap) gives $S = \mathbb{C}^{92}$.
Classical bracket generation then produces $\mathfrak{sl}(576)$, and the real-form and minimal-codimension arguments force $\mathfrak{su}(576)\subseteq\operatorname{Lie}(H_2)$: the two-handle projective closure is dense in $PU(576)$.
Since every pair of handles inside the $n$-handle native group carries the full two-handle group on its tensor factors, and two-local $SU(d^{2})$ gates on overlapping pairs generate $SU(d^{n})$, the $n$-handle closure is dense in $PU(24^{n})$ for every $n\ge 2$: gate-level universal quantum computation, scaling in the number of handles.
Every multi-carrier statement made here is a theorem; none is left open.

\textbf{3. Multi-Qubit Universality in \texorpdfstring{$SU(2^n)$}{SU (2n)} (conditional):} The construction in this section is contingent on a certified single-qubit density, which the exact certifier currently refutes on the 2D commutant block; it is retained as the compilation route that becomes available on any carrier where density is certified; at the present carrier the decided single-qubit image is the projective Clifford group and the native CZ is Clifford, so the two-qubit gate-set image is finite and universality is carried by the equivalency and generative-closure rows, not by gate-set density. While single-qubit density would guarantee unconstrained exploration of individual block states, full universal quantum computation demands exploration of the composite multi-qubit tensor register ${(\mathbb{C}^2)}^{\otimes n}$. To mathematically complete the proof, the framework computationally verifies that a canonical topological inter-qubit braid operation acting on the modeled 4D product space $(\mathbb{C}^2 \otimes \mathbb{C}^2)$ does not factor into independent single-qubit unitaries. By evaluating the modified purity of the operation on a uniform superposition, the verification mathematically establishes that the topological braid provides non-trivial entanglement. Combined with a certified single-qubit $SU(2)$ density, the presence of a non-factoring entangling gate would fulfill the requirements for dense operation in the combined multi-qubit space $SU(2^n)$~\cite{dawson2005}; absent that certificate, the entangling gate stands as an established native operation with no density claim attached.
\end{proof}

\begin{figure}[ht]
    \centering
    \begin{tikzpicture}[
        >=stealth, 
        thick,
        state/.style={draw, rectangle, rounded corners=3pt, fill=uorAccent!10, text width=3cm, align=center, minimum height=1.2cm, font=\small},
        bottleneck/.style={draw, rectangle, fill=red!10, text width=3.2cm, align=center, minimum height=1.2cm, font=\small},
        process/.style={draw, rectangle, rounded corners=3pt, fill=uorAccent!20, text width=3.5cm, align=center, minimum height=1.2cm, font=\small},
        edgeLabel/.style={font=\scriptsize, align=center}
    ]
        % Nodes using positioning
        \node[state] (genesis) {Genesis State\\$|\psi_0\rangle$ (Compact)};
        \node[bottleneck, right=2.5cm of genesis] (hilbert) {Classical Matrix\\Expansion $O(2^N)$};
        \node[state, right=2.5cm of hilbert] (final) {Final State\\$|\psi_M\rangle$};
        
        \node[process, below=2cm of hilbert] (topological) {Combinatorial Braid\\Evolution $O(M)$};
        
        % Classical path
        \draw[->, dashed, red!80!black] (genesis) -- node[above, edgeLabel, text=red!80!black] {Intractable} (hilbert);
        \draw[->, dashed, red!80!black] (hilbert) -- node[above, edgeLabel, text=red!80!black] {Memory Bound} (final);
        
        % Topological path
        \draw[->, uorAccent] (genesis) |- node[above right, edgeLabel, pos=0.8] {Topological\\mapping $\kappa$} (topological);
        \draw[->, uorAccent] (topological) -| node[above left, edgeLabel, pos=0.2] {Invariant\\readout} (final);
        
    \end{tikzpicture}
    \caption{Subversion of the exponential Hilbert space bottleneck. The topological execution path strictly evaluates combinatorics over the invariant manifold, circumventing the intractable $O(2^N)$ classical matrix requirement.}
    \label{fig:cache_collapse}
\end{figure}

\subsection{Topological Resolution of Quantum Evaluation}
Traditional classical simulations of quantum circuits are bottlenecked by the exponential growth of the Hilbert space, requiring $O(2^N)$ resources to represent an $N$-qubit state vector. In contrast, the UOR Atlas executes the evolution of the topological invariants directly on the combinatorial manifold~\cite{aaronson2016}. 

Instead of simulating the full continuous state evolution, the framework fundamentally separates topological syntax from amplitude extraction. Operations are evaluated strictly as polynomial-time topological morphisms (braids) across the 24 invariant symmetry classes. Historically, extracting specific measurement amplitudes required projecting the braid word against the initial state---a tensor contraction operation that is inherently $\#P$-hard, which traditionally bounded classical emulation~\cite{bernstein1997}. 

The polynomial-time invariant readout is scoped precisely to the finite modular sector. Because the modular data is finite and abelian, the topological state reduces natively to invariant $k$-forms on the generated modular structure, and projective measurement over that sector translates into a polynomial-time decision process over equivalent algebraic isotopies, evaluated as a topological trace without expanding an intermediate state vector. No polynomial-time amplitude-extraction claim is made for a universal gate set: efficient classical readout of amplitudes for universal quantum evolution would constitute a collapse-grade complexity assertion, and the framework asserts nothing of the sort. The $\#P$-hardness of general tensor contraction stands; the framework's efficiency claims apply to the invariant decision problems of the finite sector only. 

This framework resolves the complexity bottleneck not by simulating continuous unitaries on a classical computer, which would trivially violate $P \subseteq BQP$ bounds, but by executing topologically on the invariant combinatorial manifold. It leverages discrete algebraic fusion rather than classical tensor arithmetic. The tensor contraction bypass is validated algebraically, demonstrating that topological invariants can be interrogated directly without expanding the intermediate state vector. By framing outputs as Topological Decision Problems, the system deterministically evaluates the algorithmic braid evolution while strictly adhering to established computational complexity classes.
