\section{Formal Verification of Universal Gate Sets}

The transition from a coherent topological space to an operational quantum compiler involves the synthesis of a universal quantum gate set. Within the UOR Atlas, the foundational operations---braiding and fusion---are constructed via explicit modular $S$ and $T$ matrices modeled over the MTC framework.

\subsection{The Modular \texorpdfstring{$S$}{S} and \texorpdfstring{$T$}{T} Matrices}
In topological quantum computing, anyons are characterized by two fundamental behaviors: how they fuse together, and how they rotate around each other. These behaviors are mathematically captured by two matrices: the $S$-matrix and the $T$-matrix.

The modular $S$-matrix defines the topological S-transformation (a $\pi/2$ spatial rotation) and encodes the mutual braiding statistics of the anyon classes—essentially describing what happens when one anyon loops completely around another. It also dictates the "quantum dimensions" of the objects $d_a = S_{0a}/S_{00}$~\cite{rowell2009}, which measures how rapidly the state space grows when many anyons are present. The $T$-matrix encodes the topological twists (the phase accumulated by an anyon undergoing a $2\pi$ self-rotation, like a particle spinning in place). Together, the $S$ and $T$ matrices generate a projective representation of the modular group $SL(2, \mathbb{Z})$.

The formal validation infrastructure of this work rigorously tests the structural properties of these matrices. The computational suite programmatically asserts that the generated matrices strictly obey the required algebraic invariants:
\begin{align}
    S^4 &= I, \qquad S^2 = C \\
    {(ST)}^3 &= p^{+} S^2, \qquad p^{+} = e^{2\pi i c/8} \\
    \sum_z N_{ab}^z S_{zx} &= \frac{S_{ax}S_{bx}}{S_{0x}} \quad (\text{Verlinde formula})
\end{align}
where $C$ is the charge-conjugation permutation, $N_{ab}^z$ are the fusion coefficients~\cite{verlinde1988}, and $c$ is the topological central charge. For the Atlas-native pointed category the central charge satisfies $c \equiv 5 \pmod{8}$, so the anomaly phase $p^{+} = e^{2\pi i c/8} \neq 1$: the plain relation ${(ST)}^3 = S^2$ holds only for the anomaly-free reference theory $D(\mathbb{Z}_n)$. The verifier derives $p^{+}$ independently from the Gauss sum $(\sum_a d_a^2 \theta_a)/D$ and verifies ${(ST)}^3 = p^{+} S^2$, $S^4 = I$, and $S^2 = C$ against it.

\subsection{The Braiding \texorpdfstring{$R$}{R}-Matrix and Yang-Baxter Equations}
When two anyons are swapped in space, their physical positions are exchanged. Multi-anyon entanglement is algebraically enacted via the Braiding $R$-Matrix, which mathematically maps the fusion state of two objects into their newly braided state. 

For a topological computer to function reliably, swapping anyons must be a consistent operation regardless of the order in which adjacent swaps occur. Think of three adjacent parallel threads: swapping the left two, then the right two, then the left two again should yield the exact same braided structure as swapping the right two, then the left two, then the right two. For the topology to remain consistent, the $R$-matrix must satisfy this precise geometrical constraint, known as the Yang-Baxter equation~\cite{jimbo1989}:
\begin{equation}
    (R \otimes I)(I \otimes R)(R \otimes I) = (I \otimes R)(R \otimes I)(I \otimes R)
\end{equation}

\begin{figure}[ht]
    \centering
    \begin{tikzpicture}[scale=1.2, line width=1.5pt]
        \colorlet{c1}{uorAccent}
        \colorlet{c2}{uorDark}
        \colorlet{c3}{gray}

        % Left side: (R \otimes I)(I \otimes R)(R \otimes I)
        \draw[c3] (2,0) to[out=90,in=270] (2,1) to[out=90,in=270] (1,2) to[out=90,in=270] (0,3);
        
        \draw[white, line width=5pt] (1,0) to[out=90,in=270] (0,1) to[out=90,in=270] (0,2) to[out=90,in=270] (1,3);
        \draw[c2] (1,0) to[out=90,in=270] (0,1) to[out=90,in=270] (0,2) to[out=90,in=270] (1,3);
        
        \draw[white, line width=5pt] (0,0) to[out=90,in=270] (1,1) to[out=90,in=270] (2,2) to[out=90,in=270] (2,3);
        \draw[c1] (0,0) to[out=90,in=270] (1,1) to[out=90,in=270] (2,2) to[out=90,in=270] (2,3);
        
        \node at (3,1.5) {\Large $=$};
        
        % Right side: (I \otimes R)(R \otimes I)(I \otimes R)
        \begin{scope}[shift={(4,0)}]
        \draw[c1] (0,0) to[out=90,in=270] (0,1) to[out=90,in=270] (1,2) to[out=90,in=270] (2,3);
        
        \draw[white, line width=5pt] (1,0) to[out=90,in=270] (2,1) to[out=90,in=270] (2,2) to[out=90,in=270] (1,3);
        \draw[c2] (1,0) to[out=90,in=270] (2,1) to[out=90,in=270] (2,2) to[out=90,in=270] (1,3);
        
        \draw[white, line width=5pt] (2,0) to[out=90,in=270] (1,1) to[out=90,in=270] (0,2) to[out=90,in=270] (0,3);
        \draw[c3] (2,0) to[out=90,in=270] (1,1) to[out=90,in=270] (0,2) to[out=90,in=270] (0,3);
        \end{scope}
    \end{tikzpicture}
    \caption{The Yang-Baxter Equation representing topological equivalence of braid generators.}
    \label{fig:yang_baxter}
\end{figure}

By anchoring the definitions of these matrices in the formal mathematical schema and evaluating the formal checks, the framework provides a reproducible verification that the topological gates formulated by the Atlas are mathematically sound. Precisely, the validation suite confirms that the matrices satisfy the Yang-Baxter equation and the anomaly-corrected $SL(2, \mathbb{Z})$ relations through phase-exact (not magnitude-only) pentagon, hexagon, balancing, Verlinde, and monodromy-$S$ checks at tolerance $10^{-9}$, with the $S$ and $T$ matrices independently cross-derived in exact $\mathbb{Q}(\zeta_{24})$ arithmetic. This establishes the capability to support robust computation within definitively bounded topological constraints.

\subsection{Scope of Universality}
While the topological operations strictly satisfy the MTC axioms, it is essential to state precisely which universality is established, and on which carrier. The exact certifier of Section~3 decides the single-qubit question negatively: on the unique 2-dimensional commutant block the projective image of the coupled generators is the finite projective Clifford group of exact order 24, so density in $SU(2)$ on that block is refuted---exactly decided, not open---and no single-qubit universal gate set exists there.

What the abelian theory does establish natively is an entangling operation. Previous topological models often fell into a "theory collision" trap---attempting to synthesize an entangler by improperly hybridizing a consistent abelian model with non-abelian fusion outcomes from a separate, mathematically obstructed category. To maintain categorical rigor, the gate set must live in one coherent theory on one register. By strictly isolating two distinct qubits embedded in the anyon state space of the coherent abelian construction, the validation suite verifies that the double-braiding monodromy inherently yields a non-trivial diagonal entangling phase gate (equivalent to Controlled-$Z$) that does not factor into single-qubit unitaries, with no non-abelian hybridisation required.

However, $CZ$ is itself a Clifford gate: combined with the exactly decided finite Clifford single-qubit image, the qubit gate-set image remains finite, and no multi-qubit density claim follows via $SU(2)$. Universality is instead carried by the qudit-carrier chain of Theorem~\ref{thm:density}: density in $PU(22)$ on the 22-dimensional irreducible block, density in $PU(576)$ on the irreducible two-handle pair carrier, and density in $PU(24^{n})$ for every $n \ge 2$ by two-local composition~\cite{brylinski2002}. The established model is gate-level universal quantum computation on 24-dimensional-per-handle carriers, within a single consistent topological theory.

\begin{corollary}[Encoded-qubit universality]
\label{cor:encoded}
Let a $k$-qubit register be embedded in the $n$-handle carrier by a fixed inclusion of its $2^k$ computational basis states into the $24^n$ carrier basis, with $2^k \le 24^n$. Density of the closure in $PU(24^n)$ (Theorem~\ref{thm:density}) means every carrier unitary is approximated to arbitrary precision by a generator word; in particular each encoded single- and two-qubit gate, block-embedded as $U\oplus I$, is approximated to any $\epsilon$ by Solovay-Kitaev in $PU(d)$~\cite{dawson2005}, the residual acting as at most an $\epsilon$ leakage out of the code subspace per gate. Encoded universal qubit computation on the register therefore follows.
\end{corollary}

The machine-checked content of Corollary~\ref{cor:encoded} is exact: at $n = 2$ (carrier $24^2 = 576$) and $k = 2$, the implementation verifies over $\mathbb{Q}(\zeta_{24})$ that the encoded generator set---a Hadamard on each logical qubit and a genuine $CZ$ entangler (diagonal $(1,1,1,-1)$, which does not factor as a product of single-qubit diagonals)---is exactly unitary with its defining relations. The block map $e(U) = U\oplus I$ is a coordinate embedding, so its $*$-homomorphism property is structural (automatic) and is not dressed as a certificate; what is machine-checked over $\mathbb{Q}(\zeta_{24})$ is that the embedding is \emph{faithful and relation-preserving}: the code inclusion is an injective in-range index map, each explicit embedded operator $e(g)$ is unitary and acts as $g$ on the code block and as the identity off it, and the embedded generators satisfy the encoded relations (${e(H_i)}^2 = {e(CZ)}^2 = I$, the two $e(H_i)$ commute). The $PU(576)$ density premise is imported from the pair-carrier certificate, and the chosen code inclusion is pinned by its content address $\kappa$. Image closedness ($SU(2^{k})\oplus I$ closed in $SU(576)$) and density itself are the cited structural facts on which Corollary~\ref{cor:encoded} rests, recorded as a $\texttt{build}$ row rather than machine-checked beyond the certificate.
