\section{Complex Algorithmic Rollups and Complexity Boundaries}

Having established a coherent, albeit finite, gate set, the UOR Atlas is equipped to compile \textit{algorithmic rollups}. An algorithmic rollup is the process of taking a classical sequence of quantum gates and "rolling it up" by compiling the operations into a unified topological representation. The system dynamically maps the operational sequences into robust combinatorial sequences known as \textit{braid words}. 

\begin{figure}[h]
\centering
\begin{tikzpicture}[
    node distance=1.5cm,
    block/.style={rectangle, draw, fill=blue!10, text width=3cm, text centered, rounded corners, minimum height=1.2cm},
    arrow/.style={thick,->,>=stealth}
]
    \node[block] (algo) {Quantum Algorithm (e.g., QPE, Shor's)};
    \node[block, right=2cm of algo] (compiler) {UOR Atlas Topological Compiler};
    \node[block, right=2cm of compiler, fill=green!10] (braid) {Braid Word over $\mathcal{G}$ (Clifford $\sigma,\tau,\mu$; flow $\Phi$)};
    \node[block, below=1cm of braid, fill=red!10] (eval) {Topological Readout (Polynomial-time measurement)};

    \draw[arrow] (algo) -- (compiler) node[midway, above] {Decomposition};
    \draw[arrow] (compiler) -- (braid) node[midway, above] {Synthesis};
    \draw[arrow, dashed] (braid) -- (eval) node[midway, right] {Measurement};
\end{tikzpicture}
\caption{Algorithmic compilation pipeline. The classical quantum algorithm is decomposed and synthesized into a robust combinatorial braid word representing the invariant topological state. For the finite modular sector, extraction is framed as a discrete invariant decision problem and measurement over that sector is polynomial time; no polynomial amplitude-extraction claim is made for a universal gate set.}
\label{fig:compilation}
\end{figure}

\subsection{The Topological Compilation Functor}
\label{sec:compiler_def}
We fix the object referred to throughout as the \emph{topological compiler} and state its guarantees separately from those of its evaluation.

\begin{definition}[Gate alphabet and admissibility]
\label{def:alphabet}
The source alphabet is $\mathcal{A} = \{H, X, T, \mathrm{CNOT}\} \cup \{R_\alpha(\theta) : \alpha \in \{x,y,z\},\ \theta \in \mathbb{R}\}$. A rotation $R_\alpha(\theta)$ is \emph{admissible for a carrier} $\mathcal{K}$ when $\theta$ lies within the compiler's phase tolerance $\epsilon$ of a phase realizable on $\mathcal{K}$: on the single-handle Clifford carrier this is a discrete generator phase (the finite phase net), and on a density-certified carrier ($PU(22)$ or $PU(576)$, Theorem~\ref{thm:density}) it is any phase, realized to precision $\epsilon$ by a deterministic bounded spectral-flow search. (Solovay-Kitaev in $PU(d)$~\cite{dawson2005} attains the asymptotically optimal word length $O(\log^{c}(1/\epsilon))$; the present implementation uses a simpler one-parameter equidistribution search, whose word length is bounded by the fixed search depth for a fixed $\epsilon$.)
\end{definition}

\begin{definition}[Compilation functor $\mathcal{F}$]
\label{def:functor}
Let $\mathcal{C}$ be the free monoid of admissible circuits over $\mathcal{A}$ and $\mathcal{B}$ the free monoid of braid words over the generator alphabet $\mathcal{G}$ of the class-permutation group. The single-handle sub-alphabet $\{\sigma,\tau,\mu\}\subseteq\mathcal{G}$ is exactly the \emph{Clifford} sub-alphabet: its projective image is the order-24 Clifford group of part~(ii) of Theorem~\ref{thm:density}; the certified-carrier generators (the coupled per-handle generators and the monodromy, Section~\ref{sec:shor}) complete $\mathcal{G}$. The compiler is the monoid homomorphism $\mathcal{F}:\mathcal{C}\to\mathcal{B}$ determined by a fixed template $t:\mathcal{A}\to\mathcal{B}$ and extended by concatenation, $\mathcal{F}(g_1\cdots g_M)=t(g_1)\cdots t(g_M)$. On the discrete alphabet the templates are $t(H)=\sigma\tau\sigma$, $t(X)=\mu$, $t(T)={(\sigma\tau\tau\sigma)}^{5}$, $t(\mathrm{CNOT})=\sigma\sigma\mu\tau\sigma\tau$; on an admissible rotation, $t(R_\alpha(\theta))$ is the finite generator word returned by the carrier's synthesis routine (Section~\ref{sec:shor}).
\end{definition}

\begin{definition}[Codomain semantics]
\label{def:semantics}
The codomain carries the projective unitary action $\rho:\mathcal{B}\to PU(\dim\mathcal{K})$ of $\mathcal{G}$ on the chosen carrier $\mathcal{K}$, content-addressed by $\kappa$. On the Clifford sub-alphabet this action is exact with finite image (order 24). On a density-certified carrier the synthesis realizes an \emph{arbitrary-axis} single-qubit rotation $R_\alpha(\theta)$ ($\alpha\in\{x,y,z\}$) to within $\epsilon$: the relative-phase flow $\Phi$ provides $Z$-rotations, and the $ZXZ$ Euler decomposition $R_z(a)R_x(b)R_z(c)$ realizes any axis, with $R_x$ obtained by the Clifford Hadamard sandwich $H\,\Phi\,H$; each Euler factor is synthesized to $\epsilon/3$ as a budget, and the composite residual is then \emph{measured directly} and gated (no norm sub-additivity is assumed). Every such rotation carries a \emph{residual certificate}---the Euler flow powers and the exact (floating-point) upper bound $\lvert\Delta\rvert\ge\lVert\rho(t(R_\alpha(\theta)))-R_\alpha(\theta)\rVert_{PU}$ with $\lvert\Delta\rvert\le\epsilon$ (evaluated at the Frobenius-optimal global phase)---which is pinned by the compiled word's content address $\kappa$ and re-verified at check time, so the compilation claim and its residual are inseparable.
\end{definition}

We keep two cost measures strictly separate: \emph{compilation cost}, the work of producing $\mathcal{F}(C)$, and \emph{evaluation cost}, the work of applying $\rho(\mathcal{F}(C))$ and reading its $\kappa$-invariant. The theorem below concerns compilation cost only.

\subsection{Polynomial braid compilation}
\label{sec:compiler_thm}
\begin{theorem}[Polynomial braid compilation]
\label{thm:compile}
Fix the phase tolerance $\epsilon$. There is a constant $L(\epsilon)$---the maximal per-gate word length, equal to $20$ for the discrete templates and to the (finite, $\epsilon$-dependent) synthesis bound $B(\epsilon)$ for a certified-carrier rotation, so $L(\epsilon)=\max(20, B(\epsilon))$---such that for every admissible circuit $C$ of $M$ gates the compiler emits a braid word $\mathcal{F}(C)$ with $\lvert\mathcal{F}(C)\rvert \le L(\epsilon)\,M$, in time $O(L(\epsilon)\,M)$, and using working memory bounded by $\lvert\mathcal{F}(C)\rvert$, without ever constructing a state of the $2^n$-dimensional Hilbert space or of the $\dim\mathcal{K}$-dimensional carrier. The bound is linear in $M$ for every fixed $\epsilon$.
\end{theorem}
\begin{proof}
By Definition~\ref{def:functor}, $\mathcal{F}(C)=t(g_1)\cdots t(g_M)$ is the concatenation of the per-gate templates. Each discrete template has fixed length ($\lvert t(H)\rvert=3$, $\lvert t(X)\rvert=1$, $\lvert t(T)\rvert=20$, $\lvert t(\mathrm{CNOT})\rvert=6$), and each admissible rotation returns a finite word: on the Clifford carrier a discrete-net word of length at most $3$, and on a density-certified carrier a spectral-flow word of length at most the search bound $B(\epsilon)$, a constant for the fixed tolerance $\epsilon$. Hence every per-gate word has length at most $L(\epsilon)=\max(20, B(\epsilon))$, so $\lvert\mathcal{F}(C)\rvert=\sum_{i=1}^{M}\lvert t(g_i)\rvert\le L(\epsilon)\,M$. The homomorphism is realized by appending template symbols to a growing word, performing $O(\lvert t(g_i)\rvert)$ work per gate; it allocates no array indexed by the Hilbert dimension or the carrier dimension, so the running time is $O(\sum_i\lvert t(g_i)\rvert)=O(L(\epsilon)\,M)$ and the working memory is the emitted word itself.
\end{proof}

The evaluation of $\mathcal{F}(C)$ is a separate matter: on the finite modular sector it is the polynomial invariant-decision of Section~6, content-addressed by $\kappa$. No polynomial-time amplitude extraction is claimed for a universal gate set, and the $\#P$-hardness disclaimer of Section~3 stands in full.

\subsection{Amplitude Amplification and Grover's Search}
Grover's Algorithm provides a quadratic speedup for unstructured search problems on a quantum computer. The core of amplitude amplification relies on an oracle reflection followed by a global diffusion operator.

In the topological execution manifold, these operations are mapped into discrete topological braid generators ($\sigma, \tau, \mu$). On a density-certified carrier (Theorem~\ref{thm:density}) the continuous rotations of amplitude amplification are synthesizable natively; on the Clifford carrier the reflections stay discrete. The algorithm compiles to a deterministic sequence of braids that efficiently represents the search space on the invariant manifold. The accompanying witness is an exact-arithmetic reference evaluation of a fixed small instance: for $n = 3$ qubits with a single marked item, the marked-state probability after the optimal iteration count is computed exactly as $121/128$. This is a classical exact computation validating the algorithm mathematics; no speedup claim is attached.

\subsection{Quantum Fourier Transform (QFT) and Phase Estimation}
The Quantum Fourier Transform (QFT) relies critically on precise fractional phase rotations with exponentially decreasing phase angles ($2\pi/2^k$). 

Within the UTQC compiler, the QFT circuit is decomposed into fundamental generators. On a density-certified carrier the rotations compile \emph{unconditionally}: since the projective closure is dense there (Theorem~\ref{thm:density}, $PU(22)$ / $PU(576)$), the compiler synthesizes any fractional continuous phase rotation to precision $\epsilon$ by a deterministic bounded search over spectral-flow powers---the relative flow phase between two eigenspaces of integer eigenvalue gap $d$ is $k\,d$, and $\{k\,d \bmod 2\pi\}$ is dense (as $d/2\pi$ is irrational), so a bounded $k$ lands within $\epsilon$. Each synthesized word's phase is checked exactly against its target. On the single-handle Clifford carrier no such synthesis exists (density refuted), and the compiler admits only exact discrete phases. The accompanying witness is an exact-arithmetic reference evaluation of a fixed instance: a 3-bit Quantum Phase Estimation readout recovers the phase $1/8$ exactly by integer peak minimization, with the guarantee $|\theta - k/M| \le 1/(2M)$ verified exactly. This is a classical exact computation validating the algorithm mathematics; no speedup claim is attached.

\subsection{Shor's Algorithm and the Classical Bounding Factor}
\label{sec:shor}
The core of Shor's integer factorization lies in its period-finding subroutine, which utilizes QPE coupled with a modular exponentiation unitary.

By structurally composing the QPE rollups, Shor's period finding compiles to a topological braid word. Because Shor's QPE uses the continuous-phase QFT rotations, its compilation is unconditional exactly on a density-certified carrier: the implementation compiles each controlled-phase rotation $\pi/2^{j}$ of the $4$-bit QFT on the $22$-dimensional certified carrier, verifies the synthesized spectral-flow word realizes its target phase within $\epsilon$, and then \emph{replays} the compiled instance against the exact reference evaluation---period finding for base $2$ modulo $15$ recovers the period $4$ exactly, verified against the modular-exponentiation permutation. The exact evaluation is the semantic oracle for the result; the compilation step is what becomes unconditional on the certified carrier. These algorithm witnesses---Grover, QPE, and Shor alike---are classical exact computations over fixed small instances that validate the compiled algorithm mathematics; they carry no quantum-speedup claim, and the complexity disclaimers of Section~3 apply in full: no polynomial-time amplitude extraction is claimed for a universal gate set, and the $\#P$-hardness of general tensor contraction stands.
