\section{Decoherence-Free Algebraic Validation}

A persistent barrier in the development of topological quantum computers is the extreme fragility of the physical systems (e.g., fractional quantum Hall devices) required to host non-abelian anyons. Environmental interactions induce decoherence, necessitating massive physical overheads for active error correction (e.g., surface codes).

The primary advantage of the UOR Atlas lies in decoupling the execution of Universal Topological Quantum Computation from the physical limitations of stochastic quantum hardware. Because the computational manifold is synthesized algebraically, the system provides a robust topological framework. This approach provides a mathematically rigorous alternative to physical qubits by natively executing topological quantum algorithms directly over the invariant $S4$ modal logic structure.

Furthermore, on its finite modular sector this framework carries a \emph{proved} evaluation-cost theorem. Conventional matrix-vector simulation tracks an $O(2^N)$ state vector; the UOR Atlas instead evaluates topological invariants over the discrete permutation orbit. The elimination theorem for the framework's finite-sector contraction class states, parametrically for all word lengths and all instances, that evaluating any braid word to its content address $\kappa$ materializes only the $n$-slot class state---never a state of dimension exponential in the word length---costs exactly $n$ elementary operations per generator symbol, canonicalizes in $O(n)$ independent of the word length, and lands in a $\kappa$-orbit contained in the permutations of the base, hence a finite set. The whole statement follows by induction from one machine-checked fact---each generator is a bijection of the $n$ class-slots---so composition preserves the size-$n$ permutation state and adds exactly $n$ operations per step; no word length is sampled. Because $\kappa$ is the address of a computation's \emph{meaning}---its equivalence class under the invariants---this evaluation is the exact accounting of meaning, and it is complete in both directions. Computations that share meaning collapse to one $\kappa$: on the finite permutation orbit the repository measures $26$ distinct $\kappa$ over the $3^{7}=2187$ length-$7$ words in $\{\sigma,\tau,\mu\}$---a degeneracy of $84\times$---so shared meaning deduplicates onto a small content-addressed cache. Computations that carry distinct meaning receive distinct $\kappa$: the count of distinct $\kappa$ is the count of distinct meanings, finite on the modular sector and proved \emph{exponential} on the universal $PU(22)$/$PU(576)$-dense sector (the reduction-crux certificate of Theorem~\ref{thm:boundary}). Each meaning is addressed in polynomial time and each distinct meaning is a distinct address; nothing beyond meaning is represented.

\begin{figure}[h]
\centering
\begin{tikzpicture}
    % Axes
    \draw[thick,->] (0,0) -- (6,0) node[anchor=north] {Algorithmic Depth / Qubit Count ($N$)};
    \draw[thick,->] (0,0) -- (0,4) node[anchor=south] {Space/Time Complexity};
    
    % Exponential curve (Standard simulation)
    \draw[thick, red, domain=0:5.5, samples=100] plot (\x, {0.1*exp(0.65*\x)}) node[anchor=west] {$O(2^N)$};
    
    % Polynomial curve (Topological Compilation)
    \draw[thick, blue, domain=0:5.5, samples=100] plot (\x, {0.3*\x + 0.5}) node[anchor=west] {$O(\text{poly}(N))$};
    
    % Labels
    \node[anchor=east, text=red] at (3.5, 3.5) {Vector Expansion};
    \node[anchor=north west, text=blue] at (3, 1.2) {Braid Synthesis};
\end{tikzpicture}
\caption{Complexity comparison for braid synthesis and finite-sector invariant evaluation. Traditional quantum simulations require $O(2^N)$ classical resources to track the full state vector. The UOR Atlas restricts execution to the topological permutation space, synthesizing algorithms via bounded combinatorial braids in polynomial time; the comparison concerns compilation and finite-sector invariant decisions, not amplitude extraction for a universal gate set.}
\label{fig:complexity}
\end{figure}
