\section{Computer-Assisted Formal Verification Architecture}

The scientific value of the UOR Atlas is inextricably linked to its rigorous computer-assisted formal verification regimen. The whitepaper itself acts as a conceptual mirror to this architecture, providing the theoretical context for the executable proofs contained within the validation framework~\cite{hales2008}.

\subsection{Axiomatic Schematization}
The theoretical assertions underpinning the topological framework---including the $S4$ modal logic properties, the 24 symmetry classes, and the Mac Lane coherence axioms---are defined explicitly within a centralized, machine-readable formal schema. This conceptual model eliminates ambiguity; a topological claim is only considered valid if it corresponds to an explicitly formalized constraint.

\subsection{Automated Structural Verification}
The evaluation environment enforces these constraints through a continuous parametric assertion pipeline. 
\begin{enumerate}
    \item \textbf{Executable Mathematical Assertions:} The structural constraints are mapped directly into programmatic test suites. For instance, the topological requirement that braiding operations obey the Yang-Baxter equations is verified exhaustively across the combinatorial representation space.
    \item \textbf{Oracle Bounding:} The numerical constants and matrix coefficients utilized by the system are continuously checked against immutable external mathematical ground truths. The validation pipeline ensures the algebraic evaluation strictly matches these theoretical oracles.
    \item \textbf{Deterministic, $\kappa$-Addressed Oracles:} Every acceptance criterion references a deterministic, content-pinned reference exclusively---the exact $\mathbb{Q}(\zeta_{24})$ certifiers, the $\kappa$-addressed substrate witnesses, and the F1 constant snapshot pinned by digest and upstream tag. This is a hard requirement, not a convenience: an exactness criterion (``the computed value equals the certificate'') is only well posed against a reference that returns the same value on every run, so a nondeterministic or floating-tolerance reference is inadmissible. The paper claims and the dictionary rows inherit this discipline; the algebraic certifiers carry no $\epsilon$ and no pseudo-random state that could affect a verdict---and this is \emph{witnessed}, not merely stated: a gating source scan (row \texttt{eps-free-decision-path}) asserts the certifier's verdict path contains no floating-point token or external-entropy source outside explicitly delimited instrumentation spans (the $10^{-9}$ cross-checks against the runtime construction and a report-tier trace mirror), and that the sparse mod-$p$ projection generators are deterministic fixed-seed. The delimitation is explicit: the $10^{-9}$ comparisons are instrumentation-tier and the compiler's synthesis tolerance $\epsilon$ is a declared parameter of a \texttt{build} row, not of a verdict.
    \item \textbf{Status-Ledger Honesty:} Every row of the V\&V dictionary carries one of exactly three statuses, enforced by an honesty meta-gate: \textit{some-true} marks an externally sourced established fact (an F1 theorem, a realized uor-addr operation, or a holospaces-substrate fact) that the implementation must reproduce; \textit{build} marks a precisely scoped in-repo construction, validated against the universal MTC axioms and exact in-repo certificates; and \textit{open} marks a genuine unknown, measured and reported but never asserted true. Unfulfilled logic is therefore not prohibited by fiat but rendered unassertable: a claim without a fulfilling deterministic execution can only carry the \textit{open} status, whose contract is measurement-only.
\end{enumerate}

This verification regimen constrains the academic claims presented in this document to their recorded statuses: exactly certified facts, axiom-validated constructions, and measured observations are distinguished explicitly and never conflated.

\begin{figure}[h]
\centering
\begin{tikzpicture}[
    node distance=2cm,
    box/.style={rectangle, draw, fill=gray!10, text width=3.5cm, text centered, rounded corners, minimum height=1.5cm},
    arrow/.style={thick,->,>=stealth}
]
    \node[box] (schema) {Formal Schema\\(\textit{e.g., $S4$ Axioms, Mac Lane Coherence})};
    \node[box, right=of schema] (engine) {Topological Verification Engine};
    \node[box, right=of engine] (ci) {Automated Computational Verification};

    \draw[arrow] (schema) -- (engine) node[midway, above] {Axioms};
    \draw[arrow] (engine) -- (ci) node[midway, above] {Proofs};
\end{tikzpicture}
\caption{The computer-assisted verification architecture. Mathematical axioms are translated into a formal schema, executed as programmatic tests via the topological engine, and rigorously verified against external ground truths by an automated computational pipeline.}
\label{fig:verification}
\end{figure}
